**Density Dependence**

### 2023-03-13 Rmd/backUp/condWhb.Rmd
moving files to backup

---
title: "**Blue Whiting**"
subtitle: "Emprirical evidence for density dependence"
author: "L Kell & M Pastoors"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
mathjax: TRUE

@@

source("../R/get_dropbox.R")
dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

>>>>

---
title: "**Blue Whiting**"
subtitle: "Emprirical evidence for density dependence"
author: "L Kell & M Pastoors"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE

@@

source("../R/get_dropbox.r")
dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

### 2023-03-13 Rmd/backUp/mseMac.Rmd
moving files to backup

sch=subset(sch,year==2050)
sch=merge(sch,bmsy,by=".id")
sch=merge(sch,msy,by=".id") 
```

```{r, kobe-base, fig.height=8,fig.width=8}
p1=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="Base"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey")) 
```

```{r,kb-m, fig.height=8,fig.width=8}
p2=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","red","grey","grey")) 
```

```{r,kobe-mm, fig.height=8,fig.width=8}
p3=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey")) 
```

```{r,kobe-mmm, fig.height=8,fig.width=8}
p4=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat, M"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","grey","red")) 
```

>>>>

sch=subset(sch,year==2050)
sch=merge(sch,bmsy,by=".id")
sch=merge(sch,msy,by=".id") 
sch=merge(sch,fmsy,by=".id") 
```

```{r, kobe-base, fig.height=8,fig.width=8}
kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="Base"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey")) 
```

```{r,kb-m, fig.height=8,fig.width=8}
kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","red","grey","grey")) 
```

```{r,kobe-mm, fig.height=8,fig.width=8}
kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey")) 
```

```{r,kobe-mmm, fig.height=8,fig.width=8}
kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat, M"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","grey","red")) 
```

```{r}
source("R/kobePhaseMar4.R", echo=TRUE)

```

```{r, kobe-base-fy, fig.height=8,fig.width=8}
kobePhaseMar4(subset(transmute(subset(sch,.id=="Base"),stock=f/Fmsy,harvest=catch/msy,run=ac(f))),
                     quadcol=c("yellow","yellow","green","red"),
                     xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey"),
                     xlim=3) 
```

```{r,kb-m-fy, fig.height=8,fig.width=8}
kobePhaseMar4(subset(transmute(subset(sch,.id=="DD Mass"),stock=f/Fmsy,harvest=catch/msy,run=ac(f))),
                     quadcol=c("yellow","yellow","green","red"),
                     xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","red","grey","grey"),
                     xlim=3) 
```

```{r,kobe-mm-fy, fig.height=8,fig.width=8}
kobePhaseMar4(subset(transmute(subset(sch,.id=="DD Mass, Mat"),stock=f/Fmsy,harvest=catch/msy,run=ac(f))),
                     quadcol=c("yellow","yellow","green","red"),
                     xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey"),
                     xlim=3) 
```

```{r,kobe-mmm-fy, fig.height=8,fig.width=8}
kobePhaseMar4(subset(transmute(subset(sch,.id=="DD Mass, Mat, M"),stock=f/Fmsy,harvest=catch/msy,run=ac(f))),
                     quadcol=c("yellow","yellow","green","red"),
                     xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","grey","red"),
                     xlim=3) 
```

### 2023-03-13 Rmd/backUp/mseWhb.Rmd
moving files to backup

dat=rbind(
  cbind(What="ICES",    model.frame(FLQuants(eq,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA",     model.frame(FLQuants(eqM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA M",   model.frame(FLQuants(eqDDM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA MM",  model.frame(FLQuants(eqDDMM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA MMM", model.frame(FLQuants(eqDDMMM,rec=rec,ssb=ssb,catch=catch),drop=T)))
```

```{r, eqs}
ggplot(dat)+
  geom_line(aes(ssb,catch,col=What))
```

@@

```{r, ddM}
par["bref"]=refpts(eqM)["msy","biomass"]

for (i in dimnames(vpa)$year)
  om=ddFn(i,vpa,par,FALSE,FALSE,TRUE)
```

```{r, omCheck}
om=fwdWindow(window(vpa,end=2020),end=2080,eqM)
f =FLQuant(rep(seq(0,c(refpts(eqM)["crash","harvest"])*1.5,length.out=100),each=61),
           dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
om=fwd(om,fbar=f,sr=eqM)

oms=FLStocks("Base"     =om)
oms[["DD Mass"]]        =om
oms[["DD Mass, Mat"]]   =om
oms[["DD Mass, Mat, M"]]=om
for (iYear in ac(2020:2080)){
  oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],par,TRUE,FALSE,FALSE)
  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eqM)

  oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],par,TRUE,TRUE,FALSE)
  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eqM)

  oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],par,TRUE,TRUE,TRUE)
  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eqM)
  }
```

@@

```{r, projectons-1}
prj=FLStocks("1"=iter(oms[[1]],1:4),
             "2"=iter(oms[[2]],1:4),
             "3"=iter(oms[[3]],1:4),
             "4"=iter(oms[[4]],1:4))

F=fbar(prj[[1]][,ac(2020:2080)])%=%rep(rfpts$Fmsy,each=61)

control=as(FLQuants("f"=F),"fwdControl")
prj[[1]] =fwd(prj[[1]],control=control,sr=eqM)

for (i in ac(2020:2080)) {
    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[2]] =fwd(prj[[2]],control=control,sr=eqM)

    prj[[3]] =ddFn(i,prj[[3]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[3]] =fwd(prj[[3]],control=control,sr=eqM)

    prj[[4]] =ddFn(i,prj[[4]],par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[4]] =fwd(prj[[4]],control=control,sr=eqM)
    }
```

```{r}
plot(FLStocks(prj[[1]]),probs=c(0), worm=1:4) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming no density dependence

```{r}
names(prj)=names(oms)
plot(FLStocks(prj[[2]]),probs=c(0), worm=1:4)
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass-at-age only

```{r}
plot(FLStocks(prj[[3]]),probs=c(0), worm=1:4)
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass and maturity-at-age only

```{r}
plot(FLStocks(prj[[4]]),probs=c(0), worm=1:4)
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass, maturity and natural mortality-at-age

```{r, eval=FALSE}
ldply(prj,ldply, function(x) model.frame(FLQuants(x,biomass=function(x) biomass(x)[,"2080"],
                                                    ssb    =function(x) ssb(    x)[,"2080"],
                                                    catch  =function(x) catch(  x)[,"2080"],
                                                    f      =function(x) fbar(   x)[,"2080"]),drop=T)) 
```

```{r, projectons-2}
devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2080),,,,1]%=%0,0.3)

base=propagate(iter(prj[[1]],1),100)

@@

```{r}
plot(FLStocks(prjs[[1]])) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at Base $F_{MSY}$

```{r}
plot(FLStocks(prjs[[2]])) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, with  $F_{MSY}$ for density dependence in mass-at-age

```{r}
plot(FLStocks(prjs[[3]])) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity-at-age

```{r}
plot(FLStocks(prjs[[4]])) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity, natural mortality-at-age

```{r, MSE, eval=FALSE}
mse=hcrICES(fmsy,eql,rec(fmsy),
                      par,
                      start,   end,
                      interval,lag=lag,
                      err=err,
                      bndTac=c(0,Inf))
```

>>>>

eqCs=rbind(
  cbind(What="ICES",    model.frame(FLQuants(eq,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA",     model.frame(FLQuants(eqM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA M",   model.frame(FLQuants(eqDDM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA MM",  model.frame(FLQuants(eqDDMM,rec=rec,ssb=ssb,catch=catch),drop=T)),
  cbind(What="VPA MMM", model.frame(FLQuants(eqDDMMM,rec=rec,ssb=ssb,catch=catch),drop=T)))
```

```{r, srs}
ggplot(eqCs)+
  geom_line(aes(ssb,rec,col=What))
```

@@

```{r, ddM}
params["bref"]=refpts(eqM)["msy","biomass"]

for (i in dimnames(vpaM)$year)
  om=ddFn(i,vpaM,params,FALSE,FALSE,TRUE) 
```

```{r, omCheck}
om=fwdWindow(window(om,end=2020),end=2080,eqM)
f =FLQuant(rep(seq(0,c(refpts(eq)["crash","harvest"])*1.25,length.out=100),each=61),
           dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
om=fwd(om,fbar=f,sr=eqM)

oms=FLStocks("Base"     =om)
oms[["DD Mass"]]        =om
oms[["DD Mass, Mat"]]   =om
oms[["DD Mass, Mat, M"]]=om
for (iYear in ac(2020:2080)){
  oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],params,TRUE,FALSE,FALSE)
  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eqM)

  oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],params,TRUE,TRUE,FALSE)
  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eqM)

  oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],params,TRUE,TRUE,TRUE)
  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eqM)
  } 
```

@@

```{r, projections-1}
source("/home/laurie/Desktop/projects/ddmse/R/dd.R")

om=fwdWindow(window(iter(om,51),end=2020),end=2080,eqM)
F =propagate(fbar(om[,ac(2020:2080)]),4)
F[]=rep(rfpts$Fmsy,each=61)
control=as(FLQuants("f"=F),"fwdControl")

om =fwd(om,control=control,sr=eqM)

prj=FLStocks("Base"     =om)
prj[["DD Mass"]]        =om
prj[["DD Mass, Mat"]]   =om
prj[["DD Mass, Mat, M"]]=om
for (i in ac(2020:2080)) {
    prj[[2]] =ddFn(i,prj[[2]],params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[2]] =fwd(prj[[2]],control=control,sr=eqM)

    prj[[3]] =ddFn(i,prj[[3]],params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[3]] =fwd(prj[[3]],control=control,sr=eqM)

    prj[[4]] =ddFn(i,prj[[4]],params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
    control=as(FLQuants("f"=F[,i]),"fwdControl")
    prj[[4]] =fwd(prj[[4]],control=control,sr=eqM)
    }
```

```{r}
ts=ldply(prj, function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
                                                  f    =function(x) fbar(x), 
                                                  catch=function(x) catch(x)),drop=T))
ts=transform(ts,What=.id)
ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])

#plot(prj,probs=0,worm=seq(4))+facet_grid(qname~stock,scale="free")

ggplot()+
  geom_line( aes(ssb,catch,col=.id),   data=eqCurves)+
  geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))
```

```{r}
cast(subset(ts,year==2080),iter~What,value="catch")
```

```{r}
 ggarrange(plot(FLStocks(prj[[1]]),probs=c(0), worm=1:4)+ 
               theme(axis.ticks.y = element_blank(),
                     plot.margin = margin(r = 1),
                     strip.text.y = element_text(size=0)), 
          plot(FLStocks(prj[[2]]),probs=c(0), worm=1:4)+ 
               theme(axis.text.y = element_blank(),
                     axis.ticks.y = element_blank(),
                     axis.title.y = element_blank(),
                     plot.margin = margin(r = 1, l = 1),
                     strip.text.y = element_text(size=0)), 
          plot(FLStocks(prj[[3]]),probs=c(0), worm=1:4)+ 
               theme(axis.text.y = element_blank(),
                     axis.ticks.y = element_blank(),
                     axis.title.y = element_blank(),
                     plot.margin = margin(l = 1),
                     strip.text.y = element_text(size=0)),
          plot(FLStocks(prj[[4]]),probs=c(0), worm=1:4)+ 
               theme(axis.text.y = element_blank(),
                     axis.ticks.y = element_blank(),
                     axis.title.y = element_blank(),
                     plot.margin = margin(l = 1)  ),
          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1)) 
```

**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$

```{r, projections-2}
base=propagate(iter(fwdWindow(window(prj[[1]],end=2020),end=2080,eqM),1),100)
devRec=rlnorm(100,rec(base)[,ac(2020:2080),,,,1]%=%0,0.3)

@@

```{r, projections-3}
base=iter(fwdWindow(window(prj[[1]],end=2020),end=2080,eqM),1)

    ## No DD ####################################################
    control=as(FLQuants("f"=F),"fwdControl")
    x1     =fwd(iter(base,1),control=control,sr=eqM)

    ## With DD ####################################################
    x2=x1
    x3=x1
    x4=x1
    for (iYr in ac(2020:2080)) {
        control=as(FLQuants("f"=F[,iYr]),"fwdControl")

          ## DD M ##################################################         
          x2 =ddFn(iYr,x2,params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
          x2 =fwd(x2,control=control,sr=eqM)

          ## DD MM##################################################
          x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
          x3 =fwd(x3,control=control,sr=eqM)

          ## DD MMM ################################################
          x4 =ddFn(iYr,x4,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
          x4 =fwd(x4,control=control,sr=eqM)
          }

```{r}
smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),  
            function(y) cbind(A=x,B=y,model.frame(FLQuants(prjs[[x]][[y]][,"2040"], F=fbar,SSB=ssb,Catch=catch),drop=T))))

cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="Catch")
```

```{r}
 ggarrange(plot(prjs[[1]]) +
               theme(axis.ticks.y = element_blank(),
                     plot.margin = margin(r = 1),
                     strip.text.y = element_text(size=0)), 
          plot(prjs[[2]]) + 
               theme(axis.text.y = element_blank(),
                     axis.ticks.y = element_blank(),
                     axis.title.y = element_blank(),
                     plot.margin = margin(r = 1, l = 1),
                     strip.text.y = element_text(size=0)), 
          plot(prjs[[3]]) + 
               theme(axis.text.y = element_blank(),
                     axis.ticks.y = element_blank(),
                     axis.title.y = element_blank(),
                     plot.margin = margin(l = 1),
                     strip.text.y = element_text(size=0)),
          plot(prjs[[4]]) + 
               theme(axis.text.y = element_blank(),
                     axis.ticks.y = element_blank(),
                     axis.title.y = element_blank(),
                     plot.margin = margin(l = 1)  ),
          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1)) 
```

```{r, MSE, eval=FALSE}
mse=hcrICES(fmsy,eql,rec(fmsy),
                      params,
                      start,   end,
                      interval,lag=lag,
                      err=err,
                      bndTac=c(0,Inf))
```

### 2023-03-13 Rmd/backUp/omMac.Rmd
moving files to backup

---
title: "**Operating model**"
subtitle: "Mackerel: Operating Model with density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE

>>>>

---
title: "**Operating model**"
subtitle: "Mackerel: Operating Model with density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE

### 2023-03-13 Rmd/backUp/omWhb.Rmd
moving files to backup

---
title: "**Density Dependence**"
subtitle: "Blue whiting: Operating Model with density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE

@@

>>>>

---
title: "**Density Dependence**"
subtitle: "Blue whiting: Operating Model with density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE

@@

```{r}
save(vpaM,vpaDDM,vpaDDMM,vpaDDMMM,par,file=file.path("/home/laurie/Desktop/projects/ddmse/data/om","omWhb.RData"))
```
